---
title: "Formatting, printing and exporting tables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Formatting, printing and exporting tables}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# The difference between a dataframe and its render

Most of objects encountered throughout the **easystats** packages are "tables", i.e., a 2D matrix with columns and rows. In R, these objects are often, at their core, *dataframes*. Let's create one to use as an example:

```{r, warning=FALSE, message=FALSE}
library(insight)
library(dplyr)

df <- data.frame(
  Variable = c(1, 3, 5, 3, 1),
  Group = c("A", "A", "A", "B", "B"),
  CI = c(0.95, 0.95, 0.95, 0.95, 0.95),
  CI_low = c(3.35, 2.425, 6.213, 12.1, 1.23),
  CI_high = c(4.23, 5.31, 7.123, 13.5, 3.61),
  p = c(0.001, 0.0456, 0.45, 0.0042, 0.34)
)

df
```

When I display in in the console (calling an object - e.g. `df` - is actually equivalent to calling `print(df)`), the output looks alright, but it could be improved. Some packages, such as `knitr`, have functions to create a nicer output. For instance, in markdown, so that it can be nicely rendered in markdown documents when copied:

```{r, results='asis'}
knitr::kable(df, format="markdown")
```


Or HTML, which again makes it look great in HTML files. For instance, this code:

```{r, results='asis'}
knitr::kable(df, format="html")
```

Will be rendered like this:

```{r, echo=FALSE}
knitr::kable(df, format="html")
```

# The *insight* workflow

The **insight** package also contains function to improve the "printing", or rendering, of tables. Its design dissociates two separate and independent steps: **formatting** and **exporting**.

## Formatting

The purpose of formatting is to improve a given table, while still keeping it as a regular R dataframe, so that it can be for instance further modified by the user.


```{r}
insight::format_table(df)
```

As you can see, `format_table()` modifies columns, turning number into characters (so that it has the same amount of digits), and  detecting confidence intervals. This is usually combined with column-specific formatting functions, like `format_p()`:

```{r}
df %>% 
  mutate(p = insight::format_p(p, stars = TRUE)) %>% 
  insight::format_table()
```

TODO: What's happening here with the p-values???


## Exporting

The next step is **exporting**, which takes a dataframe and renders it in a given format, so that it looks good in the console, or in markdown, html or latex.

TODO: to complete, as it's not clear for me either!

```{r}
cat(export_table(df))
```


TODO: What about display?
